Supporting Information: Time Series Modeling of Solute Transport in an H\textsubscript{II} Phase Lyotropic Liquid Crystal Membrane

### 2019-09-18 Ben_Manuscripts/stochastic_transport/Supporting_Information.tex
Ironing out method MSDDM methodology
Justifying everything

  \clearpage
  \bibliography{transport}

>>>>

  \newpage
  \section{Estimating the Hurst Parameter}\label{section:H_estimate}

  We chose to estimate the Hurst parameter, $H$ by a least squares fit to the analytical
  autocorrelation function for fractional Brownian motion (the variance-normalized version 
  of Equation~\ref{M-eqn:fbm_autocorrelation} in the main text):

  \begin{equation}
    \gamma(k) = \dfrac{1}{2}\bigg[|k-1|^{2H} - 2|k|^{2H} + |k+1|^{2H}\bigg]
  \label{eqn:fbm_autocorrelation}
  \end{equation}

  In Figure~\ref{fig:hurst_autocorrelation}, we plotted Equation~\ref{eqn:fbm_autocorrelation}
  for different values of $H$. When $H > 0.5$, Equation~\ref{eqn:fbm_autocorrelation} decays
  slowly to zero meaning one needs to study large time lags with high frequency in order to
  obtain accurate estimates. Fortunately, all of our solutes show anti-correlated motion, so
  most of the information in Equation \ref{eqn:fbm_autocorrelation} is contained within the 
  first few lags.

  % /supporting_figures/hurst_autocorrelation.py
  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{hurst_autocorrelation.pdf}
  \caption{The analytical autocorrelation function of FBM decays to zero faster when 
  H $<$ 0.5 compared to when H $>$ 0.5.}\label{fig:hurst_autocorrelation}
  \end{figure}

  The autocovariance function of Fractional L\'evy motion is different from fractional
  Brownian motion (see Equation~\ref{M-eqn:fbm_autocorrelation}
  and~\ref{M-eqn:flm_autocovariance} of the main text), but their autocorrelation 
  structures are the same. The autocovariance function of FLM is dependent on the 
  expected value of squared draws from the underlying L\'evy distribution, $E\big[L(1)^2\big]$. 
  This is effectively the distribution's variance, which is undefined for most 
  L\'evy stable distributions due to their heavy tails. As a consequence, one should 
  expect $E\big[L(1)^2\big]$ to grow as more samples are drawn from the distribution. 
  However, we are only interested in the autocorrelation function. In order to predict
  the Hurst parameter from the autocorrelation function, we must show that it has 
  well-defined structure and is independent of the coefficient in 
  Equation~\ref{M-eqn:flm_autocovariance} of the main text. In Figure~\ref{fig:flm_autocorrelation}, 
  we plot the average autocorrelation function from an FLM process with an increasing 
  number of observations per generated sequence. For all simulations we set $H$=0.35 
  and $\alpha$=1.4. The variance-normalized autocovariance function, i.e. the autocorrelation
  function does not change with increasing sequence length. Additionally, the
  autocorrelation function of FBM, with the same $H$ is the same.

  % supporting_figures/flm_autocov.py
  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{flm_autocovariance.pdf}
  \caption{The autocorrelation function of an FLM process does not change with 
  increasing sequence length (N). It shares the same autocorrelation function as
  fractional Brownian motion (FBM). All sequenced used to make this plot were 
  generated using $H$=0.35 and $\alpha$=1.4 (for FLM).}\label{fig:flm_autocovariance}
  \end{figure}

  \section{Simulating Fractional L\'evy Motion}\label{section:sFLM}

  \subsection*{Achieving the right correlation structure}

  We simulated FLM using the algorithm of Stoev and Taqqu~\cite{stoev_simulation_2004}.
  There are no known exact methods for simulating FLM. As a consequence, passing a
  value of $H$ and $\alpha$ to the algorithm does not necessarily result in the correct
  correlation structure, although the marginal L\'evy stable distribution is correct. 
  We applied a database-based empirical correction in order to use the
  algorithm to achieve the correct marginal distribution and correlation structure.

  Stoev and Taqqu note that the transition between negatively and positively correlated
  draws occurs when $H = 1/ \alpha$. When $\alpha=2$, the marginal distribution is 
  Gaussian and $H=0.5$ as expected from FBM. We corrected the input $H$ so that the 
  value of $H$ measured based on the output sequence equaled the desired $H$. We 
  first adjusted the value of $H$ by adding ($1 / \alpha - 0.5$), effectively 
  recentering the correlation sign transition for any value of $1 \leq \alpha \leq 2$.
  This correction alone does a good job for input $H$ values near 0.5, but is
  insufficient if one desires a low values of $H$. The exact correction to $H$ is 
  not obvious so we created a database of output $H$ values tabulated as a function
  of input $H$ and $\alpha$ values. Figure~\ref{fig:hurst_correction} demonstrates the
  results of applying our correction. Without the correction, FLM realizations are
  more negatively correlated. This would result in under-predicted mean squared
  displacements when applying the model.

  % /supporting_figures/demonstrate_hurst_correction.py
  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{hurst_correction.pdf}
  \caption{Correcting the Hurst parameter input to the algorithm of Stoev and Taqqu
  results in an FLM process with a more accurate correlation structure. We generated
  sequences with an input $H$ of 0.35. We estimated $H$ by fitting the autocorrelation
  function. Without the correction, $H$ is underestimated, meaning realizations are 
  more negatively correlated than they should be.}\label{fig:hurst_correction}
  \end{figure}

  \subsection*{FLM realizations from truncated L\'evy stable distributions}

  To generate realizations from an uncorrelated truncated L\'evy process, one would
  randomly sample from the base distribution and replace values that are too large
  with new random samples from the base distribution, repeating the process until
  all samples are under the desired cut-off.

  This procedure is complicated by the correlation structure of FLM. At a high level,
  Stoev and Taqqu use Riemann-sum approximations of the stochastic integrals defining
  FLM in order to generate realizations. They do this efficiently with the help of 
  Fast Fourier Transforms. In practice, this requires one to Fourier transform a zero-padded
  vector of random samples drawn from the appropriate L\'evy stable distribution, multiply
  the vector in Fourier space by a kernel function and invert back to real space. The end
  result is a correlated vector of fractional L\'evy noise.

  If one is to truncate an FLM process, one can apply the simple procedure above for 
  drawing uncorrelated values from the marginal L\'evy stable distribution, but after
  adding correlation, the maximum drawn value is typically lower than the limit set 
  by the user. Additionally, the shape of the distribution itself changes. Analogous
  to the database used to correct the Hurst parameter, we created a database to 
  correct the input truncation parameter (the maximum desired draw). The database
  returns the value of the truncation parameter that will properly truncate the
  output marginal distribution based on $H$, $\alpha$ and $\sigma$ (the width parameter).
  Figure~\ref{fig:truncation_correction} shows the result of applying our correction.

  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{truncation_correction.pdf}
  \caption{We can accurately truncate the marginal distribution of FLM innovations by
  applying a correction to the input truncation parameter. We generated FLM sequences
  and truncated the initial L\'evy stable distribution (before Fourier transforming) at 
  a value of 0.5. After correlation structure is added, the width of the distribution 
  of fractional L\'evy noise decreases significantly. We corrected the input truncation
  parameter with our database resulting in a much more accurate distribution with 
  a maximum value close to 0.5.}\label{fig:truncation_correction}
  \end{figure}

  \clearpage
  \bibliography{stochastic_transport}
